Mixed effect tweedie model in TMB

### 2020-06-25 analysis/tmb-sensor-explore/3-fit-do.Rmd
style fit-do

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point(alpha = 0.2) +
    coord_fixed() +
    scale_color_gradient2()+
    facet_wrap(~year)
  print(g)

@@

  print("")
  print("AIC:")
  print(aic(m))

}
```

@@

d_trawl <- all_sensor %>%
  # filter missing do data 
  dplyr::filter(!is.na(do_mlpl)) %>%
  # remove 2007 data because clearly faulty or scaled differently
  dplyr::filter(year>2007) %>%
  # add day of year variable
  mutate(DOY = lubridate::yday(fe_event_start), temp = temperature_c, 
    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)) %>%
  # scale potential covariates across all surveys
  scale_predictors(predictors = c(quo(depth), quo(DOY), quo(temp)))

@@

d_trawl <- d_trawl %>% mutate(exclude = if_else(do_mlpl>8, 1, 0)) %>% 
  filter(exclude != 1)

@@

```{r}
ggplot(d_trawl, aes(X,Y, colour = do_mlpl)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(X,Y, colour = temp)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

@@

```{r}
ggplot(d_trawl, aes(depth, do_mlpl, colour = DOY, size = sqrt(do_change))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()

```

@@

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY,log_do,  colour = temp)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY,temp,  colour = temp)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4 & !(year ==2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
  geom_point() +
  facet_grid(ssid~year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, year==2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(begin=0.65)
```

```{r}
ggplot(filter(d_trawl, year==2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  #facet_wrap(~year) +
  scale_color_viridis_d(end = 0.35)
```

```{r}
ggplot(d_trawl, aes(DOY, -depth , colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
  scale_color_viridis_c() + ylab("-log(depth)")
```

@@

Remove most samples from WCVI in 2016 due to extreme values and make mesh
```{r}
d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid ==4 & DOY > 149))

@@

saveRDS(do_model_all_DOY, file = (here::here(
  paste0("analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-", 
    "800kn", #Sys.Date(), 
    ".rds")))
  )
```

@@

```{r}
predictions <- predict(do_model_all_DOY)
predictions$residuals <- residuals(do_model_all_DOY)
predictions$abs <- exp(abs(predictions$residuals))

good_rows <- filter (predictions, abs != "Inf")
sum(good_rows$abs)
19719.73-17459.49
2260.24/19719.73  # 11% reduction in abs errors
```

@@

nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year<2019)

@@

predtemp <- pred_temp %>% mutate(
  temp = est,
  temp_omega = omega_s,
  temp_epsilon = epsilon_st,
  temp_scaled = (est - model2$data$temp_mean[1])/model2$data$temp_sd[1], 
  temp_scaled2 = (temp_scaled)^2) %>% select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
predtemp <- predtemp %>% filter(year > 2007) 
predtemp$DOY_scaled <- 0

@@

# may want to filter estimates for depths less than 15 m
# pred_do <- pred_do %>% filter(depth>15)
saveRDS(pred_do, file=here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(),"-more2016.rds")))
```

@@

plot_map_raster <- function(dat, column = "est") {
    ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed()+
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

@@

Plot each survey separately
```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)

@@

pred_1 <- plot_map_raster(p_nd1, "do_est") +
  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), 
    limits = c(0, 7), 
    na.value = "gold") +
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  ggtitle("Prediction (fixed effects + all random effects)") + 
  theme(legend.position = c(.85,.25))
pred_1
```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
pred_4 <- plot_map_raster(p_nd4, "do_est") +
  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
     na.value = "red")+
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  #ggtitle(" ")+ 
  theme(legend.position = "none")
pred_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(year!=2016)
pred_16 <- plot_map_raster(p_nd16, "do_est") +
  scale_fill_viridis_c(labels = labels, breaks = c(0,2,4,6,8), limits = c(0, 7), 
     na.value = "red")+
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  #ggtitle(" ")+ 
  theme(legend.position = "none")
pred_16
```

@@

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3")#, limits = c(-0.65,1)
  # scale_fill_viridis_c()
```

```{r, eval=FALSE}
omega_s_all <- plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
   scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.65,1))+
  labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82,.75))

@@

```{r}
 plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3")
```

```{r}
p_nd1 <- p_nd %>% filter(ssid!=16) %>% filter(ssid!=4)
st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.5,0.5)) +
        labs(fill = " ") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85,.25))
st_1n3
```

```{r}
p_nd4 <- p_nd %>% filter(ssid==4) #%>% filter(year!=2016)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.5,0.5)) +
          labs(fill = "epsilon") + 
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid==16) #%>% filter(ssid!=4)
st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
  #ggtitle(" ") +
  facet_wrap(~year) +
  scale_fill_gradient2( high = "Steel Blue 4",
  low = "Red 3", limits = c(-0.5,0.5)) +
  xlim(65,345) +
  ylim(5800,6100) +
gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_16
```

@@

gridExtra::grid.arrange(
  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16 ),
  widths = c(2, 2, 1.75),
  heights = c(2, 1.5),
  layout_matrix = rbind(c(1, 2, 3),
                        c(4, 5, 6)),
  top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")
)
dev.off()
```

>>>>

  g <- ggplot(predictions, aes(X, Y, colour = residuals)) +
    geom_point(alpha = 0.2) +
    coord_fixed() +
    scale_color_gradient2() +
    facet_wrap(~year)
  print(g)

@@

  print("")
  print("AIC:")
  print(aic(m))
}
```

@@

d_trawl <- all_sensor %>%
  # filter missing do data
  dplyr::filter(!is.na(do_mlpl)) %>%
  # remove 2007 data because clearly faulty or scaled differently
  dplyr::filter(year > 2007) %>%
  # add day of year variable
  mutate(
    DOY = lubridate::yday(fe_event_start), temp = temperature_c,
    raw_depth = depth, depth = log(depth), log_do = log(do_mlpl)
  ) %>%
  # scale potential covariates across all surveys
  scale_predictors(predictors = c(quo(depth), quo(DOY), quo(temp)))

@@

d_trawl <- d_trawl %>%
  mutate(exclude = if_else(do_mlpl > 8, 1, 0)) %>%
  filter(exclude != 1)

@@

```{r}
ggplot(d_trawl, aes(X, Y, colour = do_mlpl)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

```{r}
ggplot(d_trawl, aes(X, Y, colour = temp)) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

@@

```{r}
ggplot(d_trawl, aes(depth, do_mlpl, colour = DOY, size = sqrt(do_change))) +
  geom_point() +
  facet_wrap(~year) +
  scale_color_viridis_c()
```

@@

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY, log_do, colour = temp)) +
  geom_point() +
  facet_grid(ssid ~ year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4), aes(DOY, temp, colour = temp)) +
  geom_point() +
  facet_grid(ssid ~ year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, ssid == 4 & !(year == 2016 & DOY > 149)), aes(X, Y, colour = log_do)) +
  geom_point() +
  facet_grid(ssid ~ year, scales = "free") +
  scale_color_viridis_c()
```

```{r}
ggplot(filter(d_trawl, year == 2016), aes((temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  # facet_wrap(~year) +
  scale_color_viridis_d(begin = 0.65)
```

```{r}
ggplot(filter(d_trawl, year == 2017), aes(x = (temp), log(do_mlpl), colour = as.factor(ssid))) +
  geom_point() +
  # facet_wrap(~year) +
  scale_color_viridis_d(end = 0.35)
```

```{r}
ggplot(d_trawl, aes(DOY, -depth, colour = do_mlpl)) + geom_point() + facet_wrap(~year) +
  scale_color_viridis_c() + ylab("-log(depth)")
```

@@

Remove most samples from WCVI in 2016 due to extreme values and make mesh
```{r}
d_trawl1 <- d_trawl %>% filter(!(year == 2016 & ssid == 4 & DOY > 149))

@@

saveRDS(do_model_all_DOY, file = (here::here(
  paste0(
    "analysis/tmb-sensor-explore/models/model-do-without-wcvi2016b-",
    "800kn", # Sys.Date(),
    ".rds"
  )
)))
```

@@

```{r}
predictions <- predict(do_model_all_DOY)
predictions$residuals <- residuals(do_model_all_DOY)

# predictions$abs <- exp(abs(predictions$residuals))
# pred_rows <- filter (predictions, abs != "Inf")
# sum(pred_rows$abs) # sum absolute error
# 19719.73-17459.49 # difference in absolute error between 500 and 800 knots
# 2260.24/19719.73  # 11% reduction in abs errors
```

@@

nd_all <- readRDS(here::here("analysis/VOCC/data/nd_all_synoptic.rds")) %>% filter(year < 2019)

@@

predtemp <- pred_temp %>%
  mutate(
    temp = est,
    temp_omega = omega_s,
    temp_epsilon = epsilon_st,
    temp_scaled = (est - model2$data$temp_mean[1]) / model2$data$temp_sd[1],
    temp_scaled2 = (temp_scaled)^2
  ) %>%
  select(-est, -est_non_rf, -est_rf, -omega_s, -zeta_s, -epsilon_st)
predtemp <- predtemp %>% filter(year > 2007)
predtemp$DOY_scaled <- 0

@@

saveRDS(pred_do, file = here::here(paste0("analysis/VOCC/data/predicted-DO-", Sys.Date(), "-more2016.rds")))
```

@@

plot_map_raster <- function(dat, column = "est") {
  ggplot(dat, aes_string("X", "Y", fill = column)) +
    geom_raster() +
    coord_fixed() +
    gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank())
}

@@

Plot each survey separately
```{r}
p_nd1 <- p_nd %>%
  filter(ssid != 16) %>%
  filter(ssid != 4)

@@

pred_1 <- plot_map_raster(p_nd1, "do_est") +
  scale_fill_viridis_c(
    labels = labels, breaks = c(0, 2, 4, 6, 8),
    limits = c(0, 7),
    na.value = "gold"
  ) +
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  ggtitle("Prediction (fixed effects + all random effects)") +
  theme(legend.position = c(.85, .25))
pred_1
```

```{r}
p_nd4 <- p_nd %>% filter(ssid == 4) # %>% filter(year!=2016)
pred_4 <- plot_map_raster(p_nd4, "do_est") +
  scale_fill_viridis_c(
    labels = labels, breaks = c(0, 2, 4, 6, 8), limits = c(0, 7),
    na.value = "red"
  ) +
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  # ggtitle(" ")+
  theme(legend.position = "none")
pred_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid == 16) # %>% filter(year!=2016)
pred_16 <- plot_map_raster(p_nd16, "do_est") +
  scale_fill_viridis_c(
    labels = labels, breaks = c(0, 2, 4, 6, 8), limits = c(0, 7),
    na.value = "red"
  ) +
  facet_wrap(~year) +
  labs(fill = "ml/L DO") +
  # ggtitle(" ")+
  theme(legend.position = "none")
pred_16
```

@@

```{r}
plot_map_raster(p_nd, "est_rf") +
  ggtitle("Prediction (all random effects only)") +
  facet_wrap(~year) +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3"
  ) # , limits = c(-0.65,1)
# scale_fill_viridis_c()
```

```{r, eval=FALSE}
omega_s_all <- plot_map_raster(p_nd, "omega_s") +
  ggtitle("Spatial random effects only") +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3", limits = c(-0.65, 1)
  ) +
  labs(fill = " ") +
  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.82, .75))

@@

```{r}
plot_map_raster(p_nd, "epsilon_st") +
  ggtitle("Spatiotemporal random effects only") +
  facet_wrap(~year) +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3"
  )
```

```{r}
p_nd1 <- p_nd %>%
  filter(ssid != 16) %>%
  filter(ssid != 4)
st_1n3 <- plot_map_raster(p_nd1, "epsilon_st") +
  ggtitle("Spatiotemporal random effects") +
  facet_wrap(~year) +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3", limits = c(-0.5, 0.5)
  ) +
  labs(fill = " ") +
  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = c(.85, .25))
st_1n3
```

```{r}
p_nd4 <- p_nd %>% filter(ssid == 4) # %>% filter(year!=2016)
st_4 <- plot_map_raster(p_nd4, "epsilon_st") +
  # ggtitle(" ") +
  facet_wrap(~year) +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3", limits = c(-0.5, 0.5)
  ) +
  labs(fill = "epsilon") +
  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_4
```

```{r}
p_nd16 <- p_nd %>% filter(ssid == 16) # %>% filter(ssid!=4)
st_16 <- plot_map_raster(p_nd16, "epsilon_st") +
  # ggtitle(" ") +
  facet_wrap(~year) +
  scale_fill_gradient2(
    high = "Steel Blue 4",
    low = "Red 3", limits = c(-0.5, 0.5)
  ) +
  xlim(65, 345) +
  ylim(5800, 6100) +
  gfplot::theme_pbs() + theme(axis.title.x = element_blank(), axis.title.y = element_blank(), legend.position = "none")
st_16
```

@@

gridExtra::grid.arrange(
  grobs = list(pred_1, st_1n3, omega_s_all, pred_4, st_4, st_16),
  widths = c(2, 2, 1.75),
  heights = c(2, 1.5),
  layout_matrix = rbind(
    c(1, 2, 3),
    c(4, 5, 6)
  ),
  top = grid::textGrob("Dissolved oxygen estimates from sdmTMB model")
)
dev.off()
```
